### 2024-07-29 paper.tex
chore: nuke appendix

\def\paperversiondraft{draft}
\def\paperversionnormal{normal}

% If the paper version is set to 'normal' mode keep it,
% otherwise set it to 'draft' mode.
\ifx\paperversion\paperversionnormal
\else
  \def\paperversion{draft}
\fi

% \documentclass[review, anonymous, sigplan]{acmart}
\documentclass[review, sigplan]{acmart}

\def\acmversionanonymous{anonymous}
\def\acmversionjournal{journal}
\def\acmversionnone{none}

\makeatletter
\if@ACM@anonymous
  \def\acmversion{anonymous}
\else
  \def\acmversion{journal}
\fi
\makeatother

\usepackage{colortbl}

% 'draftonly' environment
\usepackage{environ}
\ifx\paperversion\paperversiondraft
\newenvironment{draftonly}{}{}
\else
\NewEnviron{draftonly}{}
\fi

% Most PL conferences are edited by conference-publishing.com. Follow their
% advice to add the following packages.
%
% The first enables the use of UTF-8 as character encoding, which is the
% standard nowadays. The second ensures the use of font encodings that support
% accented characters etc. (Why should I use this?). The mictotype package
% enables certain features 'to­wards ty­po­graph­i­cal per­fec­tion
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{microtype}

\usepackage{xargs}
\usepackage{lipsum}
\usepackage{xparse}
\usepackage{xifthen, xstring}
\usepackage{xspace}
\usepackage{marginnote}
\usepackage{etoolbox}
\usepackage[acronym,shortcuts]{glossaries}

\input{tex/setup.tex}
\input{tex/acm.tex}

\usemintedstyle{colorful}

% Newer versions of minted require the 'customlexer' argument for custom lexers
% whereas older versions require the '-x' to be passed via the command line.
\makeatletter
\ifcsdef{minted@optlistcl@quote}
{
\newminted[mlir]{tools/MLIRLexer.py:MLIRLexerOnlyOps}{customlexer, mathescape}
\newminted[xdsl]{tools/MLIRLexer.py:MLIRLexer}{customlexer, mathescape, style=murphy}
\newminted[lean4]{tools/Lean4Lexer.py:Lean4Lexer}{customlexer, mathescape}
}
{
\newminted[mlir]{tools/MLIRLexer.py:MLIRLexerOnlyOps -x}{mathescape}
\newminted[xdsl]{tools/MLIRLexer.py:MLIRLexer -x}{mathescape, style=murphy}
\newminted[lean4]{tools/Lean4Lexer.py:Lean4Lexer -x}{mathescape}
}
\makeatother

% We use the following color scheme
% 
% This scheme is both print-friendly and colorblind safe for
% up to four colors (including the red tones makes it not
% colorblind safe any more)
%
% https://colorbrewer2.org/#type=qualitative&scheme=Paired&n=4

\definecolor{pairedNegOneLightGray}{HTML}{cacaca}
\definecolor{pairedNegTwoDarkGray}{HTML}{827b7b}
\definecolor{pairedOneLightBlue}{HTML}{a6cee3}
\definecolor{pairedTwoDarkBlue}{HTML}{1f78b4}
\definecolor{pairedThreeLightGreen}{HTML}{b2df8a}
\definecolor{pairedFourDarkGreen}{HTML}{33a02c}
\definecolor{pairedFiveLightRed}{HTML}{fb9a99}
\definecolor{pairedSixDarkRed}{HTML}{e31a1c}

\createtodoauthor{grosser}{pairedOneLightBlue}
\createtodoauthor{sid}{pairedThreeLightGreen}
\createtodoauthor{authorTwo}{pairedTwoDarkBlue}
\createtodoauthor{authorFour}{pairedFourDarkGreen}
\createtodoauthor{authorFive}{pairedFiveLightRed}
\createtodoauthor{authorSix}{pairedSixDarkRed}

\newacronym{ir}{IR}{Intermediate Representation}

\graphicspath{{./images/}}

% Define macros that are used in this paper
%
% We require all macros to end with a delimiter (by default {}) to enusure
% that LaTeX adds whitespace correctly.
\makeatletter
\newcommand\requiredelimiter[2][########]{%
  \ifdefined#2%
    \def\@temp{\def#2#1}%
    \expandafter\@temp\expandafter{#2}%
  \else
    \@latex@error{\noexpand#2undefined}\@ehc
  \fi
}
\@onlypreamble\requiredelimiter
\makeatother

\newcommand\newdelimitedcommand[2]{
\expandafter\newcommand\csname #1\endcsname{#2}
\expandafter\requiredelimiter
\csname #1 \endcsname
}

\newdelimitedcommand{toolname}{Tool}

\usepackage{booktabs}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}

\usepackage[verbose]{newunicodechar}
\newunicodechar{₁}{\ensuremath{_1}}
\newunicodechar{₂}{\ensuremath{_2}}
\newunicodechar{∀}{\ensuremath{\forall}}
\newunicodechar{α}{\ensuremath{\alpha}}
\newunicodechar{β}{\ensuremath{\beta}}

% \circled command to print a colored circle.
% \circled{1} pretty-prints "(1)"
% This is useful to refer to labels that are embedded within figures.
\DeclareRobustCommand{\circled}[2][]{%
    \ifthenelse{\isempty{#1}}%
        {\circledbase{pairedOneLightBlue}{#2}}%
        {\autoref{#1}: \hyperref[#1]{\circledbase{pairedOneLightBlue}{#2}}}%
}

% listings don't write "Listing" in autoref without this.
\providecommand*{\listingautorefname}{Listing}
\renewcommand{\sectionautorefname}{Section}
\renewcommand{\subsectionautorefname}{Section}
\renewcommand{\subsubsectionautorefname}{Section}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}

% Whitebox BitVector 
% \title{A Taxonomy of Bitvector Proof Strategies}
% Navigating the Maze of BitVectors
\title{Taming the Bitvector Bestiary}
\subtitle{A Characterization of Whitebox Bitvector Proof Strategies}

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.
% \author{First1 Last1}
% \authornote{with author1 note}          %% \authornote is optional;
% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
% \affiliation{
%   \position{Position1}
%   \department{Department1}              %% \department is recommended
%   \institution{Institution1}            %% \institution is required
%   \streetaddress{Street1 Address1}
%   \city{City1}
%   \state{State1}
%   \postcode{Post-Code1}
%   \country{Country1}
% }
% \email{first1.last1@inst1.edu}          %% \email is recommended
\author{Alex Keizer, Siddharth Bhat, Henrik B{\"o}ving, Author4, Author5, Tobias Grosser}

\begin{abstract}
\input{outline/00-abstract.tex}
\end{abstract}

% Only add ACM notes and keywords in camera ready version
% Drop citations and footnotes in draft and blind mode.
\ifx\acmversion\acmversionanonymous
\settopmatter{printacmref=false} % Removes citation information below abstract
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
\fi
\ifx\acmversion\acmversionjournal
%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
\keywords{keyword1, keyword2, keyword3} 
\fi

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\input{outline/01-introduction.tex}
\input{outline/01-API.tex}
\input{outline/01-LeanSAT.tex}
\input{outline/02-arbitrary-width.tex}
\input{outline/03-Automaton-Decision-Procedure.tex}
\input{outline/04-Non-complete-automation.tex}
\input{outline/05-Evaluation.tex}
\input{outline/06-Related-Work.tex}

\section{Conclusion}

%% Acknowledgments
% \begin{acks}                            %% acks environment is optional
%                                         %% contents suppressed with 'anonymous'
%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
%   %% acknowledge financial support and will be used by metadata
%   %% extraction tools.
%   This material is based upon work supported by the
%   \grantsponsor{GS100000001}{National Science
%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
%   conclusions or recommendations expressed in this material are those
%   of the author and do not necessarily reflect the views of the
%   National Science Foundation.
% \end{acks}

%% Bibliography
\bibliography{references}

%% Appendix
% Move \cleardouble and \appendix out of `draftonly` if you are using the
% appendix
% \begin{draftonly}
% \cleardoublepage
% \appendix
% \section{Formatting and Writing Guidelines}
% 
% These formatting guidelines aim to standardize our writing. They ensure that
% papers with multiple authors have a consistent look and that commonly occurring
% items are formatted in ways that are known to work well.
% 
% \subsection{Figures}
% \label{appendix:figures}
% 
% \paragraph{Referencing Figures} When referencing figures from the text we
% ensure the following:
% \begin{description}
%       \item [All figures are referenced] A paper with un-referenced figures
% 	      appears incomplete.
%       \item [References to figures are brief and easy to skip]~\\
%                 We minimize the number of words needed to refer to a figure. Reducing
%                 the number of non-information-carrying words directly increases
% 		the density of interesting content. When skipping references
% 		becomes easy, reading quickly while ignoring figures remains a
% 		smooth experience. The best and briefest reference to a figure
% 		is a link in parenthesis that is added after the subject
% 		representing the content depicted in a figure:\\
% 		{\color{pairedTwoDarkBlue}\textit{Figure
% 		X shows the design of A, which consists of ...}}\\
% 		$\to$ {\color{pairedFourDarkGreen}
% 		\textit{The design of A (Figure X) consists of ...}}
%       \item [The text is always self-contained without figures] ~\\ The reader
%                 should be able to read the text without ever looking at any
%                 figure. They should still understand the text and get the key
% 		message of each figure directly from the text. By not forcing
% 		the reader to analyze a figure while reading, we increase
% 		readability as the reader can continue reading without having
% 		to skip between text and figures. Such writing style also helps
% 		to guide the thoughts of the reader, who can (for a moment)
% 		trust our summary of the figure and does not need to develop
% 		their own interpretation on-the-fly, a task which often yields
% 		results that do not fit the flow of our exposition. Readers
% 		typically only feel that their reading is interrupted if there
% 		is no explanation of a figure at all. Hence, we do not need to
% 		discuss all details of a figure, but half a sentence that explains the
% 		core idea is typically sufficient for a reader to continue
% 		reading.  By making
% 		our text self-contained even when ignoring figures the reader
% 		experiences a smooth and uninterrupted reading experience.\\
% 		{\color{pairedTwoDarkBlue}
% 		\textit{The speedups are presented in Figure X. < a new topic> }}\\
% 		$\to$ {\color{pairedFourDarkGreen}\textit{Our approach outperforms the state of the art
% 		XXX-library (Figure 3) demonstrating more than 4x speedup on
% 		test case 1 and 2 and a geometric mean speedup of 1.5x over all
% 		20 test cases.}}
% \end{description}
% We reference figures in text using
% \texttt{\symbol{92}autoref\{fig:speedup\}} for a figure with label
% \texttt{fig:speedup}.  The use of autoref ensures that all references
% to figures are formatted consistently, e.g. as \autoref{fig:speedup}.
% 
% \paragraph{Color Scheme} 
% 
% In Figures we use a color scheme that is print-friendly and also visible
% with red-green blindness. The following colors are all print-friendly
% and red-green save when only using Color 1-4:
% 
% \medskip
% {
% 	\small
% \newcolumntype{a}{>{\columncolor{pairedOneLightBlue}}c}
% \newcolumntype{b}{>{\columncolor{pairedTwoDarkBlue}}c}
% \newcolumntype{d}{>{\columncolor{pairedThreeLightGreen}}c}
% \newcolumntype{e}{>{\columncolor{pairedFourDarkGreen}}c}
% \newcolumntype{f}{>{\columncolor{pairedFiveLightRed}}c}
% \newcolumntype{g}{>{\columncolor{pairedSixDarkRed}}c}
% 
% \begin{tabular}{a b d e f g}
% Color 1 & Color 2 & Color 3 & Color 4 & Color 5 & Color 6\\
% \#a6cee3 & \#1f78b4 & \#b2df8a & \#33a02c & \#fb9a99 & \#e31a1c
% \end{tabular}
% }
% 
% We de-emphasize components in figures by using additionally two shades of gray.
% Especially in complex figures, it is often helpful to de-emphasize visual
% elements that we want to represent but that should not be the focus of a
% reader's attention.
% 
% \medskip
% {
% 	\small
% \newcolumntype{h}{>{\columncolor{pairedNegOneLightGray}}c}
% \newcolumntype{i}{>{\columncolor{pairedNegTwoDarkGray}}c}
% 
% \begin{tabular}{h i}
% Color -1 & Color -2\\
% \#cacaca & \#827b7b\\
% \end{tabular}
% }
% 
% Single-color graphs are plotted in Color 1 - Light Blue.
% 
% \paragraph{Labels in Figures}
% Complex diagrams often benefit from labels inside the diagrams. We suggest to
% use a filled circle (e.g, in light blue) to highlight these numbers and use
% these references, e.g., \circled{1} implemented as \texttt{\textbackslash{}circled\{1\}}, in the text to refer to them.
% 
% \paragraph{Captions and Core Message}
% \label{appendix:captions}
% 
% Each figure should have a caption that makes a clear statement about this
% figure, as such a statement makes it easier for the reader to (in)validate the
% figure as evidence for the claim we make. Traditionally, figures often have a
% caption indicating its content:\\ {\color{pairedTwoDarkBlue} \textit{$\cdot$
% Speedup of approach A vs approach B on system X}}\\ {\color{pairedTwoDarkBlue}
% \textit{$\cdot$ Architecture diagram of our solution}}\\ While these statements
% clearly state the content of a figure at the meta-level, they often lack
% information about the precise content and the claim a figure is meant to
% evidence. While knowing that a figure is an architecture diagram is useful for
% the reader when looking at the figure the reader automatically asks two
% questions: (a) what properties set this architecture apart and (b) does its
% implementation deliver the claimed properties? Or, in more general terms, what
% claims do we aim to evidence with this figure and does the figure provide the
% needed evidence to support our claims? In theory, this information could be
% contained in the text of the paper, but to optimize for readers who skim the
% figures first, we want to offer them as part of the caption. Nevertheless, it
% makes often sense to word the caption strategically to still document the
% meta-level content of a figure.\\ $\to$ {\color{pairedFourDarkGreen}
% \textit{$\cdot$ Approach A is consistently faster than approach B, except for
% inputs that are not used in practice}}\\ $\to$ {\color{pairedFourDarkGreen}
% \textit{$\cdot$ The architecture of our design increases reusability by making
% components A, B, \& C independent of the core.}} For example, after reading the
% last caption the reader can validate if the architecture design indeed enables
% the promised independence, and we can double-check while drafting the paper that
% our figure is visualized to facilitate checking if it works as evidence. ! This
% does not mean we should mislead with our figure but rather make things easy to
% check. If our figure or data would not support our claim, it should be similarly
% easy to invalidate our claim!
% 
% \subsubsection{Plots} We use matplotlib to create performance
% plots such as \autoref{fig:speedup}. We use the following
% formatting guidelines:
% \begin{itemize}
%   \item Use a vertical y-label to make it easier to read.
%   \item Remove top and right frames to reduce visual noise
% 	and allow the reader to focus on the data in the
% 	figure.
%   \item Provide the concrete data at the top of each bar.
% \end{itemize}
% 
% \noindent
% We also suggest to follow these technical remarks:
% \begin{itemize}
%   \item Create pdf plots and do not use bitmap formats (e.g., png) to
% 	ensure high quality when zooming in.
%   \item Avoid Type-3 bitmap fonts by
% 	setting fonttype to 42.
% \end{itemize}
% 
% \begin{figure}
% \includegraphics[width=\columnwidth]{plots/speedup}
% \caption{Improved running speed after 4 weeks of training.
% }
% \label{fig:speedup}
% \end{figure}
% 
% \subsubsection{Tables} We optimize our tables for readability by removing as
% much clutter as possible, while highlighting the key structure. Markus Püschel
% (see doc/paper-writing/guide-tables.pdf) wrote a nice guide on how to make nice
% tables. \autoref{tab:simple_table} illustrates this with a simple
% table.
% 
% \begin{table}
% \ra{1.2}
% \centering
% \begin{tabular}{l l l r}
%   \toprule
%   \textbf{Animal} & \textbf{Size} & \textbf{Biotope} & \textbf{Age}\\
%   \midrule
%   Dog & Medium & Ground & 20\\
%   Cat & Medium & Ground &20 \\
%   Ant & Small & Ground & 30 \\
%   Elephant & Large & Ground & 70\\
%   Whale & Large & Water & 100\\
%   Salmon & Medium & Water & 13 \\
%   Eagle & Large & Air & 35 \\
%   \bottomrule
% \end{tabular}
% \vspace{1em}
% \caption{A table with heigh lines and emphasized header.}
% \label{tab:simple_table}
% \end{table}
% 
% \subsubsection{Listings} We aim to use minted to create listings as much as
% possible, as this allows us to edit code quickly. We use syntax highlighting
% to make the parts of the code that matter most stand out. Hence, we keep
% most code black, comments gray, and highlight just the MLIR operands that
% we care about most.
% 
% \begin{listing}[H]
% % We cannot put '{' on a line after % in draftonly mode, as the hack we used to
% % not include the draft section will interpret the listing as normal
% % latex where '%' is a comment and {} need to match, which they will
% % not if only one is commented.
% \begin{mlir}
% // This is a comment
% def @foo(%0 : !dialect.type)
% {
%   %a = dialect.op(%0) : !dialect.type // $\color{black}\circled{a}$
% }
% \end{mlir}
% \caption{A simple MLIR code example with markers. Markers can also be placed in
% 	captions and refer to labels, e.g. \circled[lst:example]{a}.}
% \label{lst:example}
% \end{listing}
% 
% \begin{listing}[H]
% \begin{lean4}
% theorem funext {f₁ f₂ : ∀ (x : α), β x}
%   (h : ∀ x, f₁ x = f₂ x) : f₁ = f₂ := by
%   show extfunApp (Quotient.mk' f₁) = 
%        extfunApp (Quotient.mk' f₂)
%   apply congrArg
%   apply Quotient.sound
%   exact h
% \end{lean4}
% \caption{A simple Lean4 code example, taken from
%   \url{https://lean-lang.org/lean4/doc/syntax\_highlight\_in\_latex.html\#example-with-minted}.}
% \end{listing}
% 
% The syntax highlighting also works for xDSL-like IRs. Notice that different
% minted styles can be used for different environments. The xDSL environment uses
% the murphy-style in this case, whereas the MLIR version applies the
% colorful-style.
% 
% \begin{xdsl}
% func.func() [sym_name = "main", function_type = !fun<[
%               !iterators.columnar_batch<!tuple<[!i64]>>
%                  ], []>] {
%   ^bb0(%0 : !iterators.columnar_batch<!tuple<[!i64]>>):
%     %t : !iterators.stream<!llvm.struct<[!i64]>> =
%       iterators.scan_columnar_batch(%0 : ...)
%     %filtered : !iterators.stream<!llvm.struct<[!i64]>> =
%           iterators.filter(%input : …) [predicateRef = @s0]
%     iterators.sink(%filtered : !iterators.stream<!tuple<[!i64]>>)
%     func.return()
% }
% 
% func.func() [sym_name = "s0", function_type = !fun<[
%     !llvm.struct<[!i64]>], [!i1]>] {
%   ^bb0(%struct : !llvm.struct<[!i64]>):
%     %id : !i64 = llvm.extractvalue(%struct : ...)
%               [position = [0 : !index]]
%     %five  : !i64 = arith.constant() [value = 5 : !i64]
%     %cmp : !i1 = arith.cmpi(%id : !i64, %five : !i64)
%               [predicate = 4 : !i64]
%     func.return(%cmp : !i1)
% }
% \end{xdsl}
% 
% \section{Writing}
% 
% A couple of hints with respect to how we write text.
% 
% \subsection{Citations}
% \label{appendix:citations}
% 
% \subsubsection{Do not use numerical citations as nouns}
% Especially when working with numerical citations (e.g., [1]) the use of
% citations as nouns reduces readability. Hence, we do not use numerical citations
% as nouns and instead expand these citations with \texttt{\textbackslash{}citet} to the
% authornames.
% \\
% {\color{pairedTwoDarkBlue}
% \textit{[1] showed that .. $\dots$}}\\
% $\to$ {\color{pairedFourDarkGreen}\textit{Author et al. [1] showed}}
% 
% \subsubsection{Prefer meaningful text over citations as textual content}
% While acknowledging authors of work is important, maximizing the amount of technical
% content (outside of a historic perspective) typically makes text more direct and
% concrete. Hence, we avoid the discussion of who did what in text if the
% historic context does not add meaning or empty words can be replaced by an immediate
% citation. E.g, in the following the words `introduced in` are not carrying
% information and can be dropped.\\
% {\color{pairedTwoDarkBlue}
% \textit{we extend PreviousIdea introduced in [1] $\dots$ by}}\\
% $\to$ {\color{pairedFourDarkGreen}\textit{we extend PreviousIdea [1] by}}
% 
% 
% \subsubsection{Managing acronyms automatically}
% Managing acronyms manually can lead to situations where the specific term is not properly expanded upon first use or when it is introduced.
% The \texttt{acronym} package is useful to avoid such situations and provides full control over acronyms.
% For example, assume we have defined an acronym with \texttt{\textbackslash{}newacronym\{ir\}\{IR\}\{Intermediate Representation\}}:
% \begin{itemize}
%   \item Upon first use of \texttt{\textbackslash{}ac\{ir\}} we get: \ac{ir}.
%   \item On the second reference: \ac{ir}.
%   \item If we need to force expansion for a figure or a background section where the term is first described, we can use \texttt{\textbackslash{}acf\{ir\}} which gives: \acf{ir}.
%   \item We can obtain plural form using \texttt{\textbackslash{}acp\{ir\}} giving: \acp{ir}.
% \end{itemize}
% 
% \end{draftonly}

\end{document}